Topological Eulerian Synthesis of Slow Motion Periodic Videos

### 2018-01-31 Paper/paper.tex
More writing

Our approach to slow motion templates is {\em Eulerian}; that is, we process the video pixel by pixel with no tracking.  Eulerian approaches for video synthesis are attractive due to their ease of implementation.  For instance, a simple Fourier bandpass filtering and amplifying pixels can successfully elucidate subtle periodic motions in videos \cite{wu2012eulerian, wadhwa2013phase}.  Additionally, time domain Eulerian approaches have been used to synthesize infinitely playing ``video textures'' using random walks \cite{schodl2000video}, or to synthesize perfectly looping video templates after grouping pixels with similar periods together and devising a common period for all groups \cite{Liao2013VideoLoops,Liao2015VideoLoops}.  There are some pitfalls with Eulerian techniques, however.  Purely Eulerian amplification techniques are known to fail for large motions, and they also require the user to manually specify frequencies of interest \cite{wu2012eulerian, wadhwa2013phase}.  Furthermore, drift from period to period due to shaking and varying motions from cycle to cycle can present a problem, as can unrelated motion in the background \cite{stauffer1999adaptive}.

To address the problems with an Eulerian representation, we devise a geometric/topological Eulerian framework byconstructing {\em sliding window embeddings} of our videos (Section~\ref{sec:slidingwindow}).  Sliding window embeddings, or ``delay reconstructions,'' have found a diverse array of applications in activity recognition \cite{frank2010activity,venkataraman2016shape}, gene expression data \cite{perea2015sw1pers}, EEG analysis \cite{stam2005nonlinear, plesnik2014detection}, audio and music analysis \cite{herzel1994analysis,serra2009cross,bello2011measuring,traliemoebius}, video analysis \cite{tralie2017quasi}, and motion capture analysis \cite{venkataraman2016shape}.  Sliding window embeddings of periodic time series, in particular, form samples of a topological loop \cite{perea2015sliding}, as do generalized multivariate sliding windows on periodic video data, regardless of the type of motion present \cite{traliehigh, tralie2017quasi}.  Furthermore, as shown by the authors of \cite{tralie2017quasi}, sliding window embeddings on Eulerian videos provide a sort of ``time regularization,'' which can mitigate the effects of drift, leading to a cleaner loop.  Our geometric framework also allows us to use topological data analysis \cite{edelsbrunner2000topological,edelsbrunner2008persistent,edelsbrunner2010computational,carlsson2009topology,ghrist2014elementary} and fundamental frequency estimation \cite{Mcleod05asmarter} to autotune the temporal and spatial scales of the periodic motion, respectively, so that no user intervention is required.  Our embedding also makes it easy to assign each window a circular phase $\phi \in [0, 2\pi]$ using Laplacian Eigenmaps \cite{belkin2003laplacian}.  We then re-order the sliding windows by phase and exploit the sliding window structure to have different windows vote on the final pixels in each frame of the template, which further mitigates drift (Section~\ref{sec:cyclereordering}).  As an illustration, Figure~\ref{fig:ConceptFigure} shows an example of our technique on a sampled periodic time series (i.e. single pixel grayscale video).  There are only 12 samples per period in the original signal, so the details of each period coarse and noisy.  However, once they are re-sorted, we get a nice, fine-detailed representation of one period.  The extension of this technique to videos will be described in detail in Section~\ref{sec:methods}.

Many prior works use a combination of sliding window embeddings and topological data analysis, such as detecting the Lorenz attractor \cite{de2012topological}, quantifying circadian rhythms in gene expression data \cite{perea2015sliding}, detecting chatter in mechanical systems \cite{khasawneh2016chatter}, detecting wheezing in audio \cite{emrani2014real}, quantifying periodic activites in motion capture data \cite{vejdemo2015cohomological, venkataraman2016persistent}, quantifying vocal fold anomalies in video \cite{tralie2017quasi}, though we believe we are the first to use a sliding window + TDA framework to {\em synthesize} a new result.  Our use of the graph Laplcian to parameterize time series is also unique, though we are inspired by works that have used it to re-arrange images around a loop as a pre-processing step for structure from motion \cite{averbuch2015ringit}, and works which use vector diffusion maps to re-order the frames of microscope images a developing embryo around a loop \cite{dsilva2015diffusionvecordering}.  We use the combination of TDA + Laplacian eigenmaps as a simple alternative to the more complicated alternative cohomology circular coordinates \cite{de2011persistent}.

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{BlockDiagram.pdf}
\caption{An block diagram of our technique, illustrated by a 1D periodic time series with additive Gaussian noise ($n_t \in \mathcal{N}(0, 1)$).  We estimate the period to be 24 (a multiple of the fundamental frequency), and perform a 24-length sliding window embedding.  From this, we use the mean of the birth and death time of the largest persistence dot to autotune the scale $\sigma$ for the graph Laplacian, from which we derive circular coordinates $\phi$, which can be used to reorder the sliding windows and use them to vote on a final slow motion template.}
\label{fig:ConceptFigure}
\end{figure*}

@@

TDA addresses problems of scale

@@

Talk about how it can pick up on harmonics (e.g. jumping jacks), which is why this is important

>>>>

Our approach to slow motion templates is {\em Eulerian}; that is, we process the video pixel by pixel with no tracking.  Eulerian approaches for video synthesis are attractive due to their simplicity and ease of implementation.  For instance, a simple Fourier bandpass filtering and amplifying pixels can successfully elucidate subtle periodic motions in videos \cite{wu2012eulerian, wadhwa2013phase}.  Additionally, time domain Eulerian approaches have been used to synthesize infinitely playing ``video textures'' using random walks \cite{schodl2000video}, or to synthesize perfectly looping video templates after grouping pixels with similar periods together and devising a common period for all groups \cite{Liao2013VideoLoops,Liao2015VideoLoops}.  There are some pitfalls with Eulerian techniques, however.  Purely Eulerian amplification techniques are known to fail for large motions, and they also require the user to manually specify frequencies of interest \cite{wu2012eulerian, wadhwa2013phase}.  Furthermore, drift from period to period due to shaking and varying motions from cycle to cycle can present a problem, as can unrelated motion in the background \cite{stauffer1999adaptive}.

To address the problems with an Eulerian representation, we devise a geometric/topological Eulerian framework byconstructing {\em sliding window embeddings} of our videos (Section~\ref{sec:slidingwindow}).  Sliding window embeddings, or ``delay reconstructions,'' have found a diverse array of applications in activity recognition \cite{frank2010activity,venkataraman2016shape}, gene expression data \cite{perea2015sw1pers}, EEG analysis \cite{stam2005nonlinear, plesnik2014detection}, audio and music analysis \cite{herzel1994analysis,serra2009cross,bello2011measuring,traliemoebius}, video analysis \cite{tralie2017quasi}, and motion capture analysis \cite{venkataraman2016shape}.  Sliding window embeddings of periodic time series, in particular, form samples of a topological loop \cite{perea2015sliding}, as do generalized multivariate sliding windows on periodic video data, regardless of the type of motion present \cite{traliehigh, tralie2017quasi}.  Furthermore, as shown by the authors of \cite{tralie2017quasi}, sliding window embeddings on Eulerian videos provide a sort of ``time regularization,'' which can mitigate the effects of drift, leading to a cleaner loop.  Our geometric framework also allows us to use topological data analysis \cite{edelsbrunner2000topological,edelsbrunner2008persistent,edelsbrunner2010computational,carlsson2009topology,ghrist2014elementary} and fundamental frequency estimation \cite{Mcleod05asmarter} to autotune the temporal and spatial scales of the periodic motion, respectively, so that no user intervention is required.  Our embedding also makes it easy to assign each window a circular phase $\phi \in [0, 2\pi]$ using Laplacian Eigenmaps \cite{belkin2003laplacian}.  We then re-order the sliding windows by $\phi$ and exploit the sliding window structure to have different windows vote on the final pixels in each frame of the template, which further mitigates drift (Section~\ref{sec:cyclereordering}).  As an illustration, Figure~\ref{fig:ConceptFigure} shows an example of our technique on a sampled periodic time series (i.e. single pixel grayscale video).  There are only 12 samples per period in the original signal, so the details of each period coarse and noisy.  However, once they are re-sorted, we get a nice, fine-detailed representation of one period.  The extension of this technique to videos will be described in detail in Section~\ref{sec:methods}.

Many prior works use a combination of sliding window embeddings and topological data analysis, such as detecting the Lorenz attractor \cite{de2012topological}, quantifying circadian rhythms in gene expression data \cite{perea2015sliding}, detecting chatter in mechanical systems \cite{khasawneh2016chatter}, detecting wheezing in audio \cite{emrani2014real}, quantifying periodic activites in motion capture data \cite{vejdemo2015cohomological, venkataraman2016persistent}, and quantifying vocal fold anomalies in video \cite{tralie2017quasi}, though we believe we are the first to use a sliding window + TDA framework to {\em synthesize} a new result.  Our use of the graph Laplcian to parameterize time series is also unique, though we are inspired by works that have used it to re-arrange images around a loop as a pre-processing step for structure from motion \cite{averbuch2015ringit}, and works which use vector diffusion maps to re-order the frames of microscope images a developing embryo around a loop \cite{dsilva2015diffusionvecordering}.  We use the combination of TDA + Laplacian eigenmaps as a simple alternative to the more complicated cohomology circular coordinates proposed by the authors of \cite{de2011persistent} and used on motion capture data \cite{vejdemo2015cohomological}.

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{BlockDiagram.pdf}
\caption{An block diagram of our technique, illustrated by a 1D periodic time series with additive Gaussian noise ($n_t \in \mathcal{N}(0, 1)$).  We estimate the period to be 24 (a multiple of the fundamental frequency), and perform a 24-length sliding window embedding.  From this, we use the mean of the birth and death time of the largest persistence dot to autotune the scale $\sigma$ for the graph Laplacian, from which we derive circular coordinates $\phi$, which can be used to reorder the sliding windows and use them to vote on a final slow motion template.  The result is shown in the bottom rightmost plot, superimposed with a ground truth period shows as an orange dotted line.}
\label{fig:ConceptFigure}
\end{figure*}

@@

Since sliding window embeddings of periodic videos lie on a topological loop (a manifold which is an ``intrinsic circle''), this means that, at an appropriate scale, a nearest neighbors graph built on the sliding window point cloud will approximately be a circle graph.  To parameterize this graph, we turn to the graph Laplacian \cite{chung1997spectral}.  For a graph $G$ with $N$ vertices and an $N \times N$ symmetric adjacency matrix $A$, the graph Laplacian is defined as
\begin{equation}
L = D-A
\end{equation}
where $D_{ii} = \sum_{j = 1}^N A_{ij}, D_{i \neq j} = 0$ is the diagonal matrix.  Now let $d_{ij}: V \rightarrow \mathbb{R}^+$ be a nonnegative scalar function on the vertices of $G$.  Then the adjacency matrix correponding to the {\em unweighted} graph Laplacain, $A_{\epsilon}^u(d)$, is

\begin{equation}
A_{\epsilon}^u(d)_{ij} = \left\{ \begin{array}{cc} 1 & i \neq j, d_{ij} \leq \epsilon \\ 0 & \text{otherwise} \end{array} \right\}
\end{equation}

For a circle graph, where the $A_{ij} = 1$ if $|i-j| = 1$, $i=1, j=N$, and $j=1, i=N$, or zero otherwise, the eigenvectors spanning the image of the graph laplacian are

\begin{equation}
v_{2k}[n] = \cos(2 \pi k n / N), v_{2k-1}[n] = \sin(2 \pi k n / N), k \geq 1
\end{equation}

%In other words, the graph Laplacian of an $N$-point circle graph recovers the $N$-dimensional discrete Fourier transform.

with corresponding eigenvalues $\lambda_k = 2 - 2 \cos(2 \pi k / N)$ (and $\lambda_0 = 0$ with corresponding eigenvector $v_0[n] = 1$).  When computing the eigenvectors of a circle graph numerically, therefore, $v_1$ and $v_2$ should be a sine and cosine, or orthogonal linear combinations therein.

TODO: Talk about circulant graphs instead

@@

Talk about how it can pick up on harmonics (e.g. jumping jacks), which is why this is important (talk about using prime 49)
